---
title: "p8105_hw5_jf3354"
author: Judy Fordjuoh
date: November 12, 2021
output: github_document
---

#Loading all the librarys/setting the themes,color schemes, and the graph output sizes 
```{r, echo = FALSE}
library(tidyverse)
library(dplyr)
library(readxl)
library(haven)
library(ggplot2)
library(patchwork)
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Question 1:
Describe the raw data. Create a city_state variable (e.g. “Baltimore, MD”) and then summarize within cities to obtain the total number of homicides and the number of unsolved homicides (those for which the disposition is “Closed without arrest” or “Open/No arrest”
```{r}
homicide_df = 
    read_csv("homicide-data.csv") %>% 
    mutate(
      city_state = str_c(city, state, sep = "_"),
      resolved = case_when(
        disposition == "Closed without arrest" ~ "unsolved",
        disposition == "Open/No arrest" ~ "unsolved",
        disposition == "Closed by arrest" ~ "solved"
        )
    ) %>% select(city_state,resolved) %>% 
  filter(city_state != "Tulsa_AL")

#finding the total number of homicides
total_hom = 
  homicide_df %>% 
  group_by(city_state) %>% 
  summarize(
    homicide_total = n(),
    homicide_unsolved = sum(resolved == "unsolved")
  )

total_hom


```
#The raw data set  contains 52,129 homicides (the rows) across 50 large cities in the U.S. collected by the Washington Post. There are 12 columns overall with variables that include uid, reported_date, victim_last, victim_first, victim_race, victim_age, victim_sex, city, state, lat, lon, and disposition. After cleaning up the data and making the city_state variable, there was only 2 variables and 52,128 observations. The data fram total_hom contains the total number of homicides and the total number of unsolved homicides in each city_state


## Question 2
#Create a tidy dataframe containing data from all participants, including the subject ID, arm, and observations over time. 1) Start with a dataframe containing all file names; 2) Iterate over file names and read in data for each subject using purrr::map and saving the result as a new variable in the dataframe, 3) Tidy the result; manipulate file names to include control arm and subject ID, make sure weekly observations are “tidy”, and do any other tidying that’s necessary

```{r}
#Loading the data from the hw5_data file/creating a list by merging the spreadsheet
hw5data = 
  tibble(
    path = list.files("hw5_data")) %>% 
  mutate(
    path = str_c("hw5_data/", path), 
    data = map(path, read_csv) 
     ) %>% 
  unnest(data) %>%
  mutate(
    path = str_replace(path, "hw5_data/", " "), #
    path = str_replace(path, ".csv", " ")
    ) %>%
  separate(path, into = c("group","Subject_ID"),sep = "_")

#str_c() combines multiple character vectors into a single character vector
#A map function is one that applies the same action/function to every element of an object
#str_replace function replaces some characters with some other characters in a string

#We have to separate the observations and the weeks into their own columns
hw5data_ =
  hw5data %>% 
  pivot_longer(
    week_1:week_8,
    names_to = "Week", 
    names_prefix = "week_", 
    values_to = "Observations"
  ) %>%
  mutate(
    group = str_replace(group, "exp", "Experimental Arm"),
    group = str_replace(group, "con", "Control Arm"),
  ) %>%
  relocate(Subject_ID)

#Make a spaghetti plot showing observations on each subject over time
hw5data_ %>% 
  mutate(Week = as.numeric(Week)) %>%
  ggplot(aes(x = Week, y = Observations, color = group)) +
  geom_path(aes(group = Subject_ID)) +
  geom_point() +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(
      title = "Observations for each Subject Over Time",
      x = "Week",
      y = "Observations"
      )
```
#
ANSWER: Generally, we see that the experimental arm has a higher/increasing trend in comparison to the control arm. As for the control arm, the observations seem to be stable throughout the study's weeks. 
#

## Question 3

#The code chunk below loads the iris dataset from the tidyverse package and introduces some missing values in each column. The purpose of this problem is to fill in those missing values.
#For numeric variables, you should fill in missing values with the mean of non-missing values. For character variables, you should fill in missing values with "virginica"
#Write a function that takes a vector as an argument; replaces missing values using the rules defined above; and returns the resulting vector. Apply this function to the columns of iris_with_missing using a map statement.

```{r}
#Loading the iris dataset so I can look at it
set.seed(10)

iris_with_missing = iris %>% 
  map_df(~replace(.x, sample(1:150, 20), NA)) %>%
  mutate(Species = as.character(Species))

#writing a function that replaces the NA's
jf_iris = function(x) {
   if (is.numeric(x)) { 
   replace_na(x, mean(x, na.rm = TRUE))
   }
   else if (is.character(x)) 
     {replace_na(x, "virginica")}
}

iris_completed = map(iris_with_missing, jf_iris)

#double checking to see if there are any NA's
is.na(iris_completed)
```

